How to Avoid OOTA Without Really Trying

### 2024-10-04 paper/ootavacuous.tex
ootavacuous: s/versus/vs./ in section headings
Reported-by: Alan Stern <stern@rowland.harvard.edu>
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>
ootavacuous: s/Summary and Conclusions/Summary and Conclusion/
Reported-by: Alan Stern <stern@rowland.harvard.edu>
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>
ootavacuous: Generalize initial definition of "semantic dependency"
Reported-by: Alan Stern <stern@rowland.harvard.edu>
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

In Listing~\ref{lst:Simple OOTA}
there is a \emph{semantic dependency} from line~1 to line~2 in both
T0 and T1.
(Roughly speaking, there is a semantic dependency from a given load
to a given store when \emph{all other things being equal, a change in the
value loaded can change the value stored or prevent the store from
occurring at all.}
Here the dependencies are trivial because the values stored simply
\emph{are} the values that were loaded.)
Since real-world CPUs cannot store something
until they have determined its value,\footnote{
	Another way of saying this is that real-world CPUs do not
	make their stores visible to other CPUs until those stores
	are no longer speculative.
	A similar restriction applies to compiler-based speculation.}
the stores on line~2 cannot take place until T0's and T1's CPUs
know what values are returned by the loads on line~1.
Thus the hardware orders these stores after their corresponding loads,
and this ordering prevents the OOTA result.

@@

\subsection{OOTA: rf versus rfe}
\label{sec:OOTA: rf versus rfe}

@@

\subsection{Relative versus Absolute Dependency}
\label{sec:Relative versus Absolute Dependency}

@@

\section{Summary and Conclusions}
\label{sec:Summary and Conclusions}

>>>>

In Listing~\ref{lst:Simple OOTA}
there is a \emph{semantic dependency} from line~1 to line~2 in both
T0 and T1.
(Roughly speaking, there is a semantic dependency from a given load to
a given store when \emph{all other things being equal, a change in the
value loaded can change the value stored, change the target address of
the store, or prevent the store from occurring at all.}
Here the dependencies are trivial because the values stored simply
\emph{are} the values that were loaded.)
Since real-world CPUs cannot store something
until they have determined its value,\footnote{
	Another way of saying this is that real-world CPUs do not
	make their stores visible to other CPUs until those stores
	are no longer speculative.
	A similar restriction applies to compiler-based speculation.}
the stores on line~2 cannot take place until T0's and T1's CPUs
know what values are returned by the loads on line~1.
Thus the hardware orders these stores after their corresponding loads,
and this ordering prevents the OOTA result.

@@

\subsection{OOTA: rf vs.\ rfe}
\label{sec:OOTA: rf vs. rfe}

@@

\subsection{Relative vs.\ Absolute Dependency}
\label{sec:Relative vs. Absolute Dependency}

@@

\section{Summary and Conclusion}
\label{sec:Summary and Conclusion}
